---
title: "Health exp and resources SG vs OECD vs non-OECD"
output: html_document
---

# OECD countries
## import 
```{r}
library(tidyverse)
import_fun<-function(csv, var){
  read_csv(csv) %>% select(Country= LOCATION, Year= TIME, {{var}} := Value) #select the relevant vars
}
  
GDP<-import_fun("OECD_healthGDP.csv",pct_GDP ) 
DR<- import_fun("OECD_Drs.csv", ratio_Dr)
NURSE<- import_fun("OECD_nurses.csv", ratio_Nurse)
BED<- import_fun("OECD_beds.csv", ratio_Bed)
ELDER<- import_fun("OECD_ElderPop.csv", pct_Elder)
```
## countries
the number of countries varies. identifiy all countries
```{r}
distinct_country<- function(df){
  df %>% distinct(Country) %>% pull()
}
GDP_country<- distinct_country(GDP) # 52 countries
DR_country<- distinct_country(DR) # 40 countries 
NURSE_country<-distinct_country(NURSE) # 42 countries
BED_country<-distinct_country(BED) ## 44 countries
ELDER_country<-distinct_country(ELDER) #56 countries

cmon_country<- list(BED_country, DR_country, GDP_country, NURSE_country, ELDER_country) %>% reduce(intersect) # 40 common countries
```
## join all dataframes
```{r}
# there are total of a 8 years and 40 countries
df_shell<-data.frame(Country=rep(cmon_country, each=10),
               Year=rep(2008:2017, length(cmon_country)))

#https://gist.github.com/austensen/cc59813d77f37e7746ed654790af8539
df<-list(df_shell, BED, DR, GDP, NURSE, ELDER) %>% reduce(left_join, by=c("Country","Year"))
```

## full names of `Country`
```{r}
library(countrycode)

df<-df %>% mutate(Country=countrycode(Country, origin = "iso3c", destination = "country.name"))
```
## OECD members
```{r}
library(rvest)
oecd<-read_html("http://www.oecd.org/about/members-and-partners")

# scrap table of members of oecd
oecd_df<-oecd %>% html_nodes("#members .country-list__content") %>% html_text() %>% as_tibble() 

oecd_df %>% tail()

oecd_df<-oecd_df %>% mutate(value= str_squish(value), # remove white space
                   value= word(value,1,2), # country is the 1st two word (e.g. united states)
                   value= str_replace(value, "Year", ""), 
                   value= str_squish(value)) %>% 
  filter(value!="Country")%>% rename(Country=value)

oecd_df
# which countries are non OECD
##  accept Czechia for Czech Republic, South Korea for Korea, Slovakia for Slovak Republic
noOECD_country<-df %>% distinct(Country) %>% anti_join(oecd_df) %>% 
  dplyr::filter(Country!=c("Czechia", "South Korea", "Slovakia")) %>% pull(Country)

df<-df %>% mutate(OECD=ifelse(Country %in% noOECD_country, "Non-OECD", "OECD"))
```

# Singapore
## Drs and Nurses 
```{r}
STAFF_sg<-read_csv("SG_DrsNurses.csv") 

DR_sg<-STAFF_sg %>% filter(profession %in% c("Doctors") & between(year, 2008,2017)) %>% select(Year=year, ratio_Dr=no_per_1000,  DR_ratio_pop=ratio_to_population)

NURSE_sg<- STAFF_sg %>% filter(profession %in% c("Nurses") & between(year, 2008,2017)) %>% select(Year=year, ratio_Nurse=no_per_1000,  Nurse_ratio_pop=ratio_to_population)
```
## GDP 
```{r}
GDP_sg<- read_csv("SG_healthGDP.csv") %>% filter(between(financial_year, 2008,2017)) %>% select(Year= financial_year, pct_GDP= percentage_gdp, everything()) 

```

## Beds 
```{r}
BED_sg<- read_csv("SG_bed(abs num).csv" ,skip = 4) %>% select(-X12) %>% 
  drop_na() %>% # there are footnotes
  mutate(Variables= str_replace(Variables, "\\*", ""), Variables=str_squish(Variables))

BED_sg %>% distinct(Variables)

# subtotal and total belong in same col, separate them 
## inspired by https://www.youtube.com/watch?v=TSG74voJQ3E 49min
BED_sg<-BED_sg %>% mutate(group= rep(1:5, each=4)) %>% select(group, everything()) %>% 
  group_by(group) %>% 
  mutate(Hospital_type= first(Variables)) %>% ungroup() %>% select(Hospital_type, everything()) %>%
  mutate(Owner= ifelse(Variables %in%c("Public", "Not-for-profit", "Private"), Variables, "Total")) %>% select(Hospital_type, Owner, everything()) %>% select(- c(group, Variables))

# wide-> long df 
BED_sg<-BED_sg %>% pivot_longer(-c("Hospital_type", "Owner"), names_to="Year", values_to="Beds") %>%   group_by(Year, Owner) %>% mutate(Beds_Year_Owner= sum(Beds)) %>% ungroup()

# beds -> bed/1000 
## import population 
pop_sg<-read_csv("SG_population.csv", skip = 4) %>% select(-X12) %>% drop_na() %>% #footnotes 
  pivot_longer(-Variables, names_to="Year", values_to="Pop") %>% select(-Variables)

BED_sg<-BED_sg %>% left_join(pop_sg, by="Year") %>% mutate(ratio_Bed= (Beds_Year_Owner/Pop)*1000)
BED_sg$Year<-as.numeric(BED_sg$Year)
```
## Elder pop
```{r}
ELDER_sg<-read_csv("SG_ElderPop.csv") %>% filter(between(year, 2008,2017)) %>% group_by(year) %>% summarise(pct_Elder=sum(value)) %>% ungroup() %>% rename(Year=year)
```

## join all sg dataframes
```{r}
df_sgDetailed<-list(BED_sg, DR_sg, GDP_sg, NURSE_sg, ELDER_sg) %>% reduce(left_join, by=c("Year")) %>% mutate(Country="Singapore", OECD="Singapore",
  across(c(ratio_Dr, ratio_Nurse), as.double))

# SG df for binding to df 
df_sg<-df_sgDetailed %>% 
  filter(Owner=="Total" & Hospital_type=="Acute Hospitals") %>%  # Total as owner and Acute Hospital as need only one type of hospital as all ratio_Bed are the same for year and owner and hospital type 
  select(colnames(df))
```


# bind OECD df and SG df
## compare resources against SG 
```{r}
df<-left_join(df, df_sg, by= "Year") %>% 
  mutate(CountryVsSG_beds= ifelse(ratio_Bed.y< ratio_Bed.x, "more beds", "less beds"), 
         CountryVsSG_drs=ifelse(ratio_Dr.y<ratio_Dr.x, "more drs", "less drs"), 
         CountryVsSG_nurses=ifelse(ratio_Nurse.y<ratio_Nurse.x, "more nurses", "less nurses")) %>% 
  select(-ends_with(".y")) %>% 
  rename_with(~str_remove(.x, "\\.x"), ends_with(".x"))
```

## binding OECD and sg dataframe
```{r}
# add above variable names  which added to df 
df_sg<-df_sg %>% mutate(CountryVsSG_beds="Singapore", CountryVsSG_drs="Singapore", CountryVsSG_nurses="Singapore")

df_all<-bind_rows(df, df_sg)
```

## income bracket
```{r}
INCOME<-readxl::read_excel("world bank income bracket.xls", skip=4) %>% select(Country=Economy, Income_group=`Income group`)%>% 
  mutate(Country=ifelse(Country=="Czech Republic", "Czechia", Country), 
         Country=ifelse(Country=="Korea, Rep.", "South Korea", Country), 
         Country= ifelse(Country=="Slovak Republic", "Slovakia", Country), 
         Country= ifelse(Country=="Russian Federation", "Russia", Country),
         Income_group= ifelse(Country=="Singapore", "Singapore", Income_group)) 

df_all<-left_join(df_all, INCOME, by="Country") 
```
## avg GDP
```{r}
df_all<-df_all %>% group_by(Country) %>% mutate(avg_GDP= mean(pct_GDP, na.rm = T)) %>% ungroup() 

# rank avg GDP by income bracket 
GDP_rankByincome<- function(...){
  df_all %>% filter(Income_group %in% c(...) ) %>% mutate(avg_GDP_rankByincome= dense_rank(avg_GDP)) %>% relocate(avg_GDP, avg_GDP_rankByincome, .after= pct_GDP)}
  
df_all<-bind_rows(
  GDP_rankByincome("High income", "Singapore"), 
  GDP_rankByincome("Upper middle income"), 
  GDP_rankByincome("Lower middle income"))
```


## all ratios as one column 
```{r}
df_all<-df_all %>% pivot_longer(starts_with("ratio"), names_to="Resource", values_to="per_1000") %>% mutate(Resource= str_remove(Resource, "ratio_")) %>% select(Country, Year, Resource, per_1000, everything())
```

# secondary dataset
```{r}
# function to create df of top 5 country for each resource + avg per_1000
top5<- function(R){
  df_all %>% select(Country, Year, Resource, per_1000) %>% 
  filter(Year==2017 , Resource==R) %>%
  slice_max(per_1000,n=5) %>% mutate(avgTop5_per1000= mean(per_1000))}

df_top5resource2017<-bind_rows(top5("Bed"), top5("Dr"), top5("Nurse"), 
# bind singapore too           
        df_all %>% select(Country, Year, Resource, per_1000) %>% mutate(avgTop5_per1000=NA) %>% filter(Year==2017, Country=="Singapore"))%>%
# arrange Singapore at the bottom for each Resource 
  group_by(Resource) %>%  arrange(avgTop5_per1000, .by_group=T) %>% ungroup() %>% 
  # imputae 
  fill(avgTop5_per1000) %>% 
  relocate(Year, Resource, everything())
```

# export 
```{r}
write_csv(df_all, path = "df_all.csv")
write_csv(df_top5resource2017, path="df_top5resource2017.csv")
knitr::knit_exit()
```

# Extra Data Variables
1. outcome measure, life expectancy at 65 years old `years_aft65`. 
In contrast to life expectancy at birth, this is one of the few healthcare outcome measures biased towards measuring the success of a caring for a geriatric population which is the theme of the presentation. 
Source file for OECD, non-OECD: OECD_LifeExp65.csv
Source file for SG: SG_LifeExp65.csv

2. export `df_sgDetailed`

3i. average year to year difference of GDP 
E.g. Spending in 2007 is 4% of GDP, spending in 2008 is 6% of GDP, spending in 2009 is 7%. Year to year difference for 2007-2008, will be 2% and year to year difference for 2008-2009 will be 1%. The average year to year difference will be 1.5%.
3ii. average year to year difference in life expectancy 
3iii. average year to year difference in elder population
3iv. absolute change in healthcare spending (as measured as percentage of GDP) in 2017 compared to 2008.
3v. rankings for the above

wrangling steps 
create a `lead` -> take the difference -> take the mean of the difference 


